---
title: "animate_covid_timeseries"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I want to try out the [gganimate package!](https://gganimate.com/index.html) So I thought I'd try it out on some covid time series data from New Zealand. I found some timeseries data here: https://github.com/CSSEGISandData/COVID-19/blob/master/who_covid_19_situation_reports/who_covid_19_sit_rep_time_series/who_covid_19_sit_rep_time_series.csv

## Get data

```{bash, eval = FALSE}
git clone https://github.com/CSSEGISandData/COVID-19.git
```

*Note. I set the above code chunk to "eval = FALSE" and manually run it once, to avoid the data being downloaded each time this document is knitted*

## Load libraries

```{r}
library(ggplot2)
library(gganimate)
library(dplyr)
library(stringr)
library(lubridate)
library(gifski)
```

## Play around with data!

Load data

```{r}
data <- read.csv("./COVID-19/who_covid_19_situation_reports/who_covid_19_sit_rep_time_series/who_covid_19_sit_rep_time_series.csv", header = TRUE, stringsAsFactors = FALSE)
```

Format columns (X has automatically been prepended, but using "check.names = FALSE" causes even more problems)

```{r}
names(data) <- sapply(str_remove_all(colnames(data),"X"),"[")
```

Extract NZ data

```{r}
data_nz <- data %>%
  filter(Country.Region == "New Zealand")
```

Manipulate data

```{r}
# Remove redundant columns
data_nz <- data_nz %>%
  select (-c(Province.States, WHO.region, WHO.region.label))

# Reformat the data for downstream plot
data_nz <- data_nz %>% tidyr::gather(date, "cases", na.rm = TRUE)

# Remove first row which is old column names after gather
data_nz <- data_nz %>% slice(-1)

# Format date column
data_nz$date <- data_nz$date %>%
  as.Date(format = "%m.%d.%y")

# Format case number column
data_nz$cases <- as.integer(data_nz$cases)
```

## Plot!

```{r}
plot <- ggplot(data_nz, aes(x = date, y = cases)) +
  geom_histogram(stat = "identity") +
  theme_linedraw() +
  labs(title = "Date: {frame_time}", y = "Number of cases") +
  transition_time(date) +
  ease_aes('linear') +
  shadow_wake(.3)

animate(plot, duration = 10, fps = 20, width = 300, height = 300, renderer = gifski_renderer())

anim_save("covid_timeseries_NZ.gif", animation = last_animation(), path = ".")
```

## R info

```{r}
R.Version()
sessionInfo()
```

